Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-09-25 paper.qmd
Updated Figure 15 with only one projection

## First projection
projection <- cbind(
    c(0.24156,0.85058,0.11564,-0.03255,-0.28106,0.27463,-0.10539,0.08429,0.13107),
    c(-0.67212,0.02745,-0.26718,-0.33126,-0.34919,0.20839,-0.14498,-0.37708,0.15455))

@@

axes_obj <- gen_axes(
  proj = projection,
  limits = 1,
  axis_pos_x = -0.4,
  axis_pos_y = -0.4,
  axis_labels = names(scaled_pbmc_data),
  threshold = 0.057)

pbmc_proj_umap_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c",
    linewidth = 1) +
  geom_point(
    #size = 0.5,
    alpha = 0.2) +
  scale_color_manual(values = c("#000000", "#33a02c")) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.6, 1)) +
  ylim(c(-0.6, 1)) +
  interior_annotation("a2", cex = 2)

## Second projection
projection <- cbind(
    c(0.00067,0.53963,-0.16241,-0.09777,0.32766,0.44543,-0.51359,-0.29884,0.00951),
    c(0.62399,0.26016,0.06441,-0.34686,-0.40466,0.24643,0.23537,0.07972,0.34402))

projection_scaled <- projection * 1.25

@@

axes_obj <- gen_axes(
  proj = projection,
  limits = 1,
  axis_pos_x = -0.6,
  axis_pos_y = -0.7,
  axis_labels = names(scaled_pbmc_data),
  threshold = 0.086)

pbmc_proj_umap_model2 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c",
    linewidth = 1) +
  geom_point(
    #size = 0.5,
    alpha = 0.2) +
  scale_color_manual(values = c("#000000", "#33a02c")) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.9, 0.5)) +
  ylim(c(-0.9, 0.5)) +
  interior_annotation("a3", cex = 2)

@@

axes_obj <- gen_axes(
  proj = projection,
  limits = 1,
  axis_pos_x = -0.4,
  axis_pos_y = -0.4,
  axis_labels = names(scaled_pbmc_data),
  threshold = 0.057)

pbmc_proj_umap_model1_best <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c",
    linewidth = 1) +
  geom_point(
    #size = 0.5,
    alpha = 0.2) +
  scale_color_manual(values = c("#000000", "#33a02c")) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.6, 1)) +
  ylim(c(-0.6, 1)) +
  interior_annotation("b2", cex = 2)

## Second projection
projection <- cbind(
    c(0.00067,0.53963,-0.16241,-0.09777,0.32766,0.44543,-0.51359,-0.29884,0.00951),
    c(0.62399,0.26016,0.06441,-0.34686,-0.40466,0.24643,0.23537,0.07972,0.34402))

projection_scaled <- projection * 1.25

@@

axes_obj <- gen_axes(
  proj = projection,
  limits = 1,
  axis_pos_x = -0.6,
  axis_pos_y = -0.7,
  axis_labels = names(scaled_pbmc_data),
  threshold = 0.086)

pbmc_proj_umap_model2_best <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c") +
  geom_point(
    #size = 0.5,
    alpha = 0.2) +
  scale_color_manual(values = c("#000000", "#33a02c")) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.9, 0.5)) +
  ylim(c(-0.9, 0.5)) +
  interior_annotation("b3", cex = 2)

>>>>

## First projection
projection <- cbind(
    c(-0.4545,-0.5533,-0.1235,0.3984,-0.2065,-0.3586,-0.0088,0.3572,-0.1368),
    c(0.4576,-0.3157,0.4708,0.1410,-0.4479,-0.1685,0.3668,-0.2692,0.1332))

@@

axes_obj <- gen_axes(
  proj = projection,
  limits = 1,
  axis_pos_x = -0.48,
  axis_pos_y = -0.48,
  axis_labels = names(scaled_pbmc_data),
  threshold = 0.082)

pbmc_proj_umap_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c") +
  geom_point(
    #size = 0.5,
    alpha = 0.2) +
  scale_color_manual(values = c("#000000", "#33a02c")) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.65, 0.65)) +
  ylim(c(-0.65, 0.65)) +
  interior_annotation("a2", cex = 2)

## Second projection
projection <- cbind(
    c(-0.4545,-0.5533,-0.1235,0.3984,-0.2065,-0.3586,-0.0088,0.3572,-0.1368),
    c(0.4576,-0.3157,0.4708,0.1410,-0.4479,-0.1685,0.3668,-0.2692,0.1332))

@@

axes_obj <- gen_axes(
  proj = projection,
  limits = 1,
  axis_pos_x = -0.48,
  axis_pos_y = -0.48,
  axis_labels = names(scaled_pbmc_data),
  threshold = 0.082)

# pbmc_proj_umap_model2 <- projected_df |>
#   ggplot(
#     aes(
#       x = proj1, 
#       y = proj2)) +
#   geom_segment(
#     data = model_df, 
#     aes(
#       x = proj1_from, 
#       y = proj2_from, 
#       xend = proj1_to, 
#       yend = proj2_to), 
#     color = "#33a02c",
#     linewidth = 1) +
#   geom_point(
#     #size = 0.5,
#     alpha = 0.2) +
#   scale_color_manual(values = c("#000000", "#33a02c")) +
#   geom_segment(
#     data=axes, 
#     aes(x=x1, y=y1, xend=x2, yend=y2), 
#     colour="grey70") +
#   geom_text(
#     data=axes,
#     aes(x=x2, y=y2), 
#     label=rownames(axes),
#     colour="grey50",
#     size = 3) +
#   geom_path(
#     data=circle, 
#     aes(x=c1, y=c2), colour="grey70") +
#   coord_fixed() +
#   xlim(c(-0.9, 0.5)) +
#   ylim(c(-0.9, 0.5)) +
#   interior_annotation("a3", cex = 2)

pbmc_proj_umap_model2 <- ggplot() +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c") +
  scale_color_manual(values = c("#000000", "#33a02c")) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.65, 0.65)) +
  ylim(c(-0.65, 0.65)) +
  interior_annotation("a3", cex = 2)

@@

axes_obj <- gen_axes(
  proj = projection,
  limits = 1,
  axis_pos_x = -0.48,
  axis_pos_y = -0.48,
  axis_labels = names(scaled_pbmc_data),
  threshold = 0.082)

pbmc_proj_umap_model1_best <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c") +
  geom_point(
    #size = 0.5,
    alpha = 0.2) +
  scale_color_manual(values = c("#000000", "#33a02c")) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.65, 0.65)) +
  ylim(c(-0.65, 0.65)) +
  interior_annotation("b2", cex = 2)

## Second projection
projection <- cbind(
    c(-0.4545,-0.5533,-0.1235,0.3984,-0.2065,-0.3586,-0.0088,0.3572,-0.1368),
    c(0.4576,-0.3157,0.4708,0.1410,-0.4479,-0.1685,0.3668,-0.2692,0.1332))

@@

axes_obj <- gen_axes(
  proj = projection,
  limits = 1,
  axis_pos_x = -0.48,
  axis_pos_y = -0.48,
  axis_labels = names(scaled_pbmc_data),
  threshold = 0.082)

# pbmc_proj_umap_model2_best <- projected_df |>
#   ggplot(
#     aes(
#       x = proj1, 
#       y = proj2)) +
#   geom_segment(
#     data = model_df, 
#     aes(
#       x = proj1_from, 
#       y = proj2_from, 
#       xend = proj1_to, 
#       yend = proj2_to), 
#     color = "#33a02c") +
#   geom_point(
#     #size = 0.5,
#     alpha = 0.2) +
#   scale_color_manual(values = c("#000000", "#33a02c")) +
#   geom_segment(
#     data=axes, 
#     aes(x=x1, y=y1, xend=x2, yend=y2), 
#     colour="grey70") +
#   geom_text(
#     data=axes,
#     aes(x=x2, y=y2), 
#     label=rownames(axes),
#     colour="grey50",
#     size = 3) +
#   geom_path(
#     data=circle, 
#     aes(x=c1, y=c2), colour="grey70") +
#   coord_fixed() +
#   xlim(c(-0.9, 0.5)) +
#   ylim(c(-0.9, 0.5)) +
#   interior_annotation("b3", cex = 2)

pbmc_proj_umap_model2_best <- ggplot() +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c") +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.65, 0.65)) +
  ylim(c(-0.65, 0.65)) +
  interior_annotation("b3", cex = 2)
